\bf A class of measurement error models for singe-cell CRISPR screens

### 2022-01-04 text/jasa/jasa.tex
fix reproducibility

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{setspace}
\usepackage{amsfonts}
\usepackage{comment}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{tabularx}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{array}
\usepackage{titlesec} 
\usepackage{xcolor}
\usepackage{float}
\usepackage[toc,page]{appendix}
\usepackage{algpseudocode}
\usepackage{afterpage}
\usepackage[
backend=biber,
style=numeric-comp,
maxcitenames=1,
maxbibnames=3,
backref=true,
doi=false,
isbn=false,
url=false,
eprint=false
]{biblatex}

@@

% JOURNAL SETTINGS
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{1}
\addtolength{\oddsidemargin}{-.5in}
\addtolength{\evensidemargin}{-1in}
\addtolength{\textwidth}{1in}
\addtolength{\textheight}{1.7in}
\addtolength{\topmargin}{-1in}
\date{\vspace{-5ex}}

\begin{document}
\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}
\if1\blind
{
  \title{\bf A new class of measurement error models for CRISPR genome engineering and single-cell RNA sequencing}
  \author{Timothy Barry$^1$, Eugene Katsevich$^2$, Kathyrn Roeder$^1$ \\
    $^1$Department of Statistics and Data Science, Carnegie Mellon University \\
    $^2$Department of Statistics and Data Science, Wharton School}
  \maketitle
} \fi

\if0\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE \bf A new class of measurement error models for CRISPR genome engineering and single-cell RNA sequencing}
\end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
CRISPR genome engineering and single-cell RNA sequencing have transformed biological discovery. Single-cell CRISPR screens unite these two technologies, linking genetic perturbations in individual cells to changes in gene expression and illuminating regulatory networks underlying diseases. Despite their promise, single-cell CRISPR screens present substantial statistical challenges. We demonstrate through theoretical and real data analyses that a standard method for estimation and inference in single-cell CRISPR screens --``thresholded regression'' -- exhibits attenuation bias and a bias-variance tradeoff as a function of an intrinsic and challenging-to-select tuning parameter. To overcome these difficulties, we introduce GLM-EIV (``GLM-based errors-in-variables''), a new method for single-cell CRISPR screen analysis. GLM-EIV extends the classical errors-in-variables model to responses and noisy predictors that are (i) exponential family-distributed and (ii) potentially impacted by the same set of confounding variables. We develop a computational infrastructure to deploy GLM-EIV across tens or hundreds of nodes on clouds (e.g., Microsoft Azure) and high-performance clusters. Leveraging this infrastructure, we apply GLM-EIV to analyze two recent, large-scale, single-cell CRISPR screen datasets, demonstrating improved performance in challenging problem settings.
\end{abstract}

\noindent
{\it Keywords:}  exponential family, errors-in-variables, GLM, CRISPR screen, differential expression
\spacingset{1.9}
\section{Introduction}
CRISPR is a genome engineering tool that has enabled scientists to precisely edit human and nonhuman genomes, opening the door to new medical therapies \cite{Rothgangl2021,Musunuru2021} and transforming basic biology research \cite{Przybyla2021}. Recently, scientists have paired CRISPR genome engineering with single-cell sequencing \cite{Dixit2016,Datlinger2017}. The resulting assays, known as a ``single-cell CRISPR screens,'' link genetic perturbations in individual cells to changes in gene expression, illuminating regulatory networks underlying human diseases and other traits \cite{Morris2021a}.

Despite their promise, single-cell CRISPR screens present substantial statistical challenges. A major difficulty is that CRISPR perturbations are assigned stochastically to cells and cannot be observed directly. As a consequence, one cannot know with certainty which cells were perturbed. Instead, one must leverage an indirect, noisy proxy of perturbation presence or absence -- namely, transcribed guide RNA counts -- to ``guess'' which cells were perturbed. Using these imputed perturbation assignments, one can attempt to estimate the effect of the perturbation on gene expression. The standard approach, which we call ``thresholded regression'' or the ``thresholding method,'' is to assign perturbation identities to cells by simply thresholding the guide RNA counts.

We study estimation and inference in single-cell CRISPR screens from a statistical perspective, formulating the data generating mechanism using a new class of measurement error models. We assume that the response variable $y$ is a GLM of an underlying predictor variable $x^*$ and vector of confounders $z$. We do not observe $x^*$ directly; rather, we observe a noisy version $x$ of $x^*$ that itself is a GLM of $x^*$ and $z$. The goal of the analysis is to estimate the effect of $x^*$ on $y$ using the observed data $(x, y, z)$ only. In the context of the biological application, $x^*$, $x$, $y$, and $z$ are CRISPR perturbations, guide RNA counts, gene expressions, and technical confounders, respectively.

Our work makes two main contributions. First, we conduct a detailed study of the thresholding method. Notably, we demonstrate on real data that the thresholding method exhibits attenuation bias and a bias-variance tradeoff as a function of the selected threshold, and we recover these phenomena in precise mathematical terms in a simplified Gaussian setting. Second, we introduce a new method, GLM-EIV (``GLM-based errors-in-variables''), for single-cell CRISPR screen analysis. GLM-EIV extends the classical errors-in-variables model to response distributions and sources of measurement error that are (i) exponential family-distributed and (ii) potentially impacted by the same set of confounding variables. GLM-EIV thereby implicitly estimates the probability that each cell was perturbed, obviating the need to explicitly impute perturbation assignments via thresholding or another heuristic. % Theoretical analyses and simulation studies indicate that GLM-EIV outperforms the thresholding method in large regions of the parameter space.
We implement several statistical accelerations (that possibly are of independent utility) to bring the cost of GLM-EIV down to within about an order of magnitude of the thresholding method.

@@

\subsection{Related work}
Motivated by the challenges of single-cell data, several authors recently have extended statistical models that (implicitly or explicitly) assume Gaussianity and homoscedasticity to a broader class of exponential family distributions. For example, Lin et al.\ \cite{Lin2021} and Townes et al.\ \cite{Townes2019}  (separately) developed eSVD and GLM-PCA, generalizations of SVD and PCA, respectively, to exponential family response distributions. Unlike their vanilla counterparts, eSVD and GLM-PCA can model gene expression counts directly, thereby capturing the relationship between the mean and variance of a gene's expression level \cite{Lause2021} and improving performance on dimension reduction tasks. We see our work as a continuation of this effort to ``port'' common statistical methods and models to single-cell count data. Our focus, however, is on regression rather than dimension reduction. We extend the classical errors-in-variables model in several directions, most importantly generalizing to exponential family-distributed response distributions and sources of measurement error.

The closest parallels to our work in the statistical methodology literature are Gr\"{u}n \& Leisch \cite{Grun2008} and Ibrahim \cite{Ibrahim1990}. Gr\"{u}n \& Leisch derived a method for estimation and inference in a $k$-component mixture of GLMs. While we prefer to view GLM-EIV as a generalized errors-in-variables method,  the GLM-EIV model is equivalent to a two-component mixture of \textit{products} of GLM densities. Ibrahim proposed a procedure for fitting GLMs in the presence of missing-at-random covariates. Our method, by contrast, involves fitting two conditionally independent GLMs in the presence of a totally latent covariate. Thus, while Ibrahim and Gr\"{u}n \& Leisch are helpful references, our estimation and inference tasks are more complex than theirs. Other relevant references include Aigner \cite{Aigner1973} and Savoca \cite{Savoca2000}.

The genomics literature has produced several applied methods for single-cell CRISPR screen analysis. Barry et al.\ developed SCEPTRE \cite{Barry2021}, a custom implementation of the conditional randomization test \cite{Candes2018, Liu2021} tailored to single-cell CRISPR screen data. SCEPTRE tests whether a given perturbation is associated with the change in expression of a given gene, adjusting for sources of confounding and ensuring robustness to expression model misspecification. In this work we tackle a set of analysis challenges that are complimentary to those addressed by SCEPTRE. Most importantly, we account for the fact that the perturbation is measured with noise, an issue that SCEPTRE circumvents via thresholding. Additionally, we seek to \textit{estimate} (with confidence) the effect size of a perturbation on gene expression change, an objective that is challenging to attain within the nonparametric hypothesis testing framework of SCEPTRE.

@@

There are several broad classes of single-cell CRISPR screen assays, each suited to answer a different set of biological questions \cite{Gasperini2019,Datlinger2021,Mimitou2019}. In this work we focus on high-multiplicity of infection (MOI) single-cell CRISPR screens. We expect the ideas that we develop for this assay to apply (with some effort) to other classes of single-cell CRISPR screens as well. In this section we motivate high MOI single-cell screens, overview the experimental protocol, and present relevant analysis challenges.

The human genome consists of genes, enhancers (segments of DNA that regulate the expression of one or more genes), and other genomic elements (that are not of importance to the current work). Genome-wide association studies (GWAS) have revealed that the majority ($>90\%$) of variants associated with diseases lie outside genes and (very likely) inside enhancers \cite{Gallagher2018}. These noncoding variants are thought to contribute to disease by modulating the expression of one or more disease-relevant genes. Scientists do not know the gene (or genes) through which most noncoding variants exert their effect, limiting the interpretability of GWAS results. A central open challenge in genetics, therefore, is to link enhancers that harbor GWAS variants to the genes that they target at genome-wide scale \cite{Gasperini2020}.

@@

\subsection{Analysis challenges}
High MOI single-cell CRISPR screens present several statistical challenges, four of which we highlight here. Throughout, we consider a single perturbation-gene pair. First, the ``treatment'' variable -- i.e., the presence or absence of a perturbation -- cannot be directly observed. Instead, perturbed cells transcribe molecules called  \textit{guide RNAs} (or \textit{gRNAs}) that serve as indirect proxies of perturbation presence. We must leverage these gRNAs to impute (explicitly or implicitly) perturbation assignments onto the cells (Figure \ref{analysis_challenges}b). Second, ``technical factors'' -- sources of variation that are experimental rather than biological in origin -- impact the measurement of both gene and gRNA expressions and therefore act as confounders (Figure \ref{analysis_challenges}b). Third, the gene and gRNA data are sparse, discrete counts. Consequently, classical statistical approaches that assume Gaussianity or homoscedasticity are inapplicable. Finally, and most subtly, sequenced gRNAs sometimes map to cells that have not received a perturbation. This phenomenon, which we ``background contamination,'' results from errors in the sequencing and alignment processes \cite{Replogle2020}. The marginal distribution of the gRNA counts is best conceptualized as a mixture model (Figure \ref{analysis_challenges}c; Gaussian distributions used for illustration purposes only). Unperturbed and perturbed cells both exhibit nonzero gRNA count distributions, but this distribution is shifted upward for perturbed cells. Figure \ref{analysis_challenges}d shows example data on four (of possibly tens or hundreds of thousands of) cells. The analysis objective is to leverage the gene expressions and gRNA counts to estimate the effect of the (latent) perturbation on gene expression, accounting for the technical factors.

In this work we analyze two large-scale, high MOI, single-cell CRISPR screen datasets published by Gasperini et al.\ \cite{Gasperini2019} and Xie et al.\ \cite{Xie2019a}. Gasperini (resp., Xie) targeted approximately 6,000 (resp., 500) candidate enhancers in a population of approximately 200,000 (resp., 100,000) cells. Gasperini additionally designed several hundred positive control, gene-targeting perturbations and 50 non-targeting, negative control perturbations to assess method sensitivity and specificity.

@@

In this work we introduced GLM-EIV (``GLM-based errors in variables''), a new method for single-cell CRISPR screen analysis. GLM-EIV extends the classical errors-in-variables model to response distributions and sources of measurement error that are exponential family-distributed and potentially impacted by the same set of confounding variables. These extensions enable GLM-EIV to resolve novel analysis challenges posed by single-cell CRISPR screens. We demonstrated through simulation studies, real data analyses, and theory that GLM-EIV outperforms thresholded regression by a considerable margin in high background contamination settings. GLM-EIV intuitively achieves this performance gain by leveraging information from \textit{both} modalities (rather than the gRNA modality alone) to assign perturbation identities to cells. On the other hand, in low background contamination settings GLM-EIV and thresholded regression work best in conjunction, with GLM-EIV providing a statistically principled and empirically effective procedure for selecting the threshold. GLM-EIV thereby neutralizes a tuning parameter that, until this point, has been selected using heuristic procedures, with little confidence that the choice is near optimal. Figure \ref{fig:dec_tree} summarizes how we anticipate GLM-EIV being used in practice.

@@

\clearpage
\section*{Figures}
% 1. Challenges
\begin{figure}[h!]
	\centering
	\includegraphics[width=1\linewidth]{../../figures/analysis_challenges/plot.pdf}
	\caption{\textbf{Experimental design and analysis challenges}: \textbf{a,} Experimental design. For a given perturbation (e.g., the perturbation indicated in blue), we partition the cells into two groups: perturbed and unperturbed. Next, for a given gene, we conduct a differential expression analysis across the two groups, yielding an estimate of the impact of the given perturbation on the given gene. \textbf{b,} DAG representing all variables in the system. The perturbation (latent) impacts both gene expression and gRNA expression; technical factors act as confounders, also impacting gene and gRNA expression. The target of estimation is the effect of the perturbation on gene expression. \textbf{c,} Schematic illustrating the ``background read'' phenomenon. Due to errors in the sequencing and alignment processes, unperturbed cells exhibit a nonzero gRNA count distribution (bottom). The target of estimation is the change in mean gene expression in response to the perturbation (top). \textbf{d}, Example data on four cells for a given perturbation-gene pair. Note that (i) the perturbation is unobserved, and (ii) the gene and gRNA data are discrete counts.}
	\label{analysis_challenges}
\end{figure}

@@

\clearpage
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.7\linewidth]{../../figures/thresholding_theoretical/plot}
	\caption{\textbf{Thresholding method bias-variance decomposition.} Bias decreases and variance increases as the threshold tends to infinity. $\beta^g_1 = 1, \beta^m_1 = 1,$ and $\pi = 0.1$ in this plot.}
	\label{thresholding_theoretical}
\end{figure}

\clearpage
% Theory
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.9\linewidth]{../../figures/main_text_sim/plot}
	\caption{\textbf{Simulation study}. Columns correspond to distributions (Poisson, NB with known $\theta$, NB with estimated $\theta$), and rows correspond to metrics (bias, MSE, coverage, CI width, and time). Methods are shown in different colors; GLM-EIV (red) is masked by accelerated GLM-EIV (blue) in several panels. GLM-EIV demonstrated superior statistical performance to the thresholding method on all metrics (rows 1-4). Accelerated GLM-EIV had substantially lower computational cost than ``vanilla'' GLM-EIV (bottom row) despite demonstrating identical statistical performance (rows 1-4).}
	\label{main_text_sim}
\end{figure}

>>>>

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{setspace}
\usepackage{amsfonts}
\usepackage{comment}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{tabularx}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{array}
\usepackage{titlesec} 
\usepackage{xcolor}
\usepackage{float}
\usepackage[toc,page]{appendix}
\usepackage{algpseudocode}
\usepackage{afterpage}
\usepackage[
backend=biber,
style=chicago-authordate,
maxcitenames=1,
maxbibnames=3,
backref=true,
doi=false,
isbn=false,
url=false,
eprint=false
]{biblatex}

@@

% JOURNAL SETTINGS
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{1}
\addtolength{\oddsidemargin}{-.5in}
\addtolength{\evensidemargin}{-1in}
\addtolength{\textwidth}{1in}
\addtolength{\textheight}{1.7in}
\addtolength{\topmargin}{-1in}

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}
\if1\blind
{
	%  \title{\bf FDR control following filtering for hierarchically structured hypotheses}
	\title{\bf A class of measurement error models for singe-cell CRISPR screens}
	%	Eugene Katsevich, Chiara Sabatti, Marina Bogomolov}
	\author{Timothy Barry\thanks{ FUNDING INFORMATION GOES HERE.}\hspace{.2cm}\\
		Department of Statistics and Data Science,\\Carnegie Mellon University\\
		and\\
		Eugene Katsevich\\
		Department of Statistics and Data Science,\\Wharton School, University of Pennsylvania\\
		and\\
		Kathryn Roeder\\
		Department of Statistics and Data Science,\\Computational Biology Department,\\ 
		Carnegie Mellon University}
	\maketitle
} \fi

\if0\blind
{
	\bigskip
	\bigskip
	\bigskip
	\begin{center}
		{\LARGE\bf Title}
	\end{center}
	\medskip
} \fi

\bigskip
\begin{abstract}
CRISPR genome engineering and single-cell RNA sequencing have transformed biological discovery. Single-cell CRISPR screens unite these two technologies, linking genetic perturbations in individual cells to changes in gene expression and illuminating regulatory networks underlying diseases. Despite their promise, single-cell CRISPR screens present substantial statistical challenges. We demonstrate through theoretical and real data analyses that a standard method for estimation and inference in single-cell CRISPR screens --``thresholded regression'' -- exhibits attenuation bias and a bias-variance tradeoff as a function of an intrinsic, challenging-to-select tuning parameter. To overcome these difficulties, we introduce GLM-EIV (``GLM-based errors-in-variables''), a new method for single-cell CRISPR screen analysis. GLM-EIV extends the classical errors-in-variables model to responses and noisy predictors that are (i) exponential family-distributed and (ii) potentially impacted by the same set of confounding variables. We develop a computational infrastructure to deploy GLM-EIV across tens or hundreds of nodes on clouds (e.g., Microsoft Azure) and high-performance clusters. Leveraging this infrastructure, we apply GLM-EIV to analyze two recent, large-scale, single-cell CRISPR screen datasets, demonstrating improved performance in challenging problem settings.
\end{abstract}

\noindent%
{\it Keywords:} exponential family, GLM, latent variable model, RNA-seq, genome engineering, cloud computing

\spacingset{1.9}

\section{Introduction}
CRISPR is a genome engineering tool that has enabled scientists to precisely edit human and nonhuman genomes, opening the door to new medical therapies \parencite{Rothgangl2021,Musunuru2021} and transforming basic biology research \parencite{Przybyla2021}. Recently, scientists have paired CRISPR genome engineering with single-cell RNA sequencing \parencite{Dixit2016,Datlinger2017}. The resulting assays, known as a ``single-cell CRISPR screens,'' link genetic perturbations in individual cells to changes in gene expression, illuminating regulatory networks underlying human diseases and other traits \parencite{Morris2021a}.

Despite their promise, single-cell CRISPR screens present substantial statistical challenges. A major difficulty is that the ``treatment'' -- i.e., the presence of absence of a CRISPR perturbation -- is assigned randomly to cells and is not directly observable. As a consequence, one cannot know with certainty which cells were perturbed. Instead, one must leverage an indirect, quantitative proxy of perturbation presence or absence to ``guess'' which cells received a perturbation. This indirect proxy takes the form of so-called guide RNA counts, with higher counts indicating that a cell is more likely to have been perturbed. The standard approach to single-cell CRISPR screen analysis is to impute perturbation assignments onto the cells by simply thresholding the gRNA counts; using these imputations, one can attempt to estimate the effect of the perturbation on gene expression. We call this standard approach ``thresholded regression'' or the ``thresholding method.''

We study estimation and inference in single-cell CRISPR screens from a statistical perspective, formulating the data generating mechanism using a new class of measurement error models. We assume that the response variable $y$ is a GLM of an underlying predictor variable $x^*$ and vector of confounders $z$. We do not observe $x^*$ directly; rather, we observe a noisy version $x$ of $x^*$ that itself is a GLM of $x^*$ and the same set of confounders $z$. The goal of the analysis is to estimate the effect of $x^*$ on $y$ using the observed data $(x, y, z)$ only. In the context of the biological application, $x^*$, $x$, $y$, and $z$ are CRISPR perturbations, guide RNA counts, gene expressions, and technical confounders, respectively.

Our work makes two main contributions. First, we conduct a detailed study of the thresholding method. Notably, we demonstrate on real data that the thresholding method exhibits attenuation bias and a bias-variance tradeoff as a function of the selected threshold, and we recover these phenomena in precise mathematical terms in a simplified Gaussian setting. Second, we introduce a new method, GLM-EIV (``GLM-based errors-in-variables''), for single-cell CRISPR screen analysis. GLM-EIV extends the classical errors-in-variables model to responses and noisy predictors that are (i) exponential family-distributed and (ii) potentially impacted by the same set of confounding variables. GLM-EIV thereby implicitly estimates the probability that each cell was perturbed, obviating the need to explicitly impute perturbation assignments via thresholding or another heuristic. We implement several statistical accelerations (that possibly are of independent utility) to bring the cost of GLM-EIV down to within about an order of magnitude of the thresholding method.

@@

\subsection{Related work}
Motivated by the challenges of single-cell data, several authors recently have extended statistical models that (implicitly or explicitly) assume Gaussianity and homoscedasticity to a broader class of exponential family distributions. For example, Lin et al.\ \parencite{Lin2021} and Townes et al.\ \parencite{Townes2019}  (separately) developed eSVD and GLM-PCA, generalizations of SVD and PCA, respectively, to exponential family response distributions. Unlike their vanilla counterparts, eSVD and GLM-PCA can model gene expression counts directly, thereby capturing the relationship between the mean and variance of a gene's expression level \parencite{Lause2021} and improving performance on dimension reduction tasks. We see our work as a continuation of this effort to ``port'' common statistical methods and models to single-cell count data. Our focus, however, is on regression rather than dimension reduction. We extend the classical errors-in-variables model in several directions, most importantly generalizing to responses and noisy predictors that are exponential family-distributed.

The closest parallels to our work in the statistical methodology literature are Gr\"{u}n \& Leisch \parencite{Grun2008} and Ibrahim \parencite{Ibrahim1990}. Gr\"{u}n \& Leisch derived a method for estimation and inference in a $k$-component mixture of GLMs. While we prefer to view GLM-EIV as a generalized errors-in-variables method,  the GLM-EIV model is equivalent to a two-component mixture of \textit{products} of GLM densities. Ibrahim proposed a procedure for fitting GLMs in the presence of missing-at-random covariates. Our method, by contrast, involves fitting two conditionally independent GLMs in the presence of a totally latent covariate. Thus, while Ibrahim and Gr\"{u}n \& Leisch are helpful references, our estimation and inference tasks are more complex than theirs. Next, we focus in this work on noisy \textit{binary} predictors rather than noisy \textit{continuous} predictors, the latter of which are more commonly used in measurement error models. \parencite{Aigner1973} and \parencite{Savoca2000} also developed measurement error models for noisy binary predictors, albeit in a different statistical context than ours.

The genomics literature has produced a few applied methods for single-cell CRISPR screen analysis. Most relevantly, Barry et al.\ developed SCEPTRE, a custom implementation of the conditional randomization test \parencite{Candes2018, Liu2021} tailored to single-cell CRISPR screen data \parencite{Barry2021}. SCEPTRE tests whether a given perturbation is associated with the change in expression of a given gene, adjusting for sources of confounding and ensuring robustness to expression model misspecification. In this work we tackle a set of analysis challenges that are complimentary to those addressed by SCEPTRE. Most importantly, we account for the fact that the perturbation is measured with noise, an issue that SCEPTRE circumvents via thresholding. Additionally, we seek to \textit{estimate} (with confidence) the effect size of a perturbation on gene expression change, an objective that is challenging to attain within the nonparametric hypothesis testing framework of SCEPTRE.

@@

There are several broad classes of single-cell CRISPR screen assays, each suited to answer a different set of biological questions \parencite{Gasperini2019,Datlinger2021,Mimitou2019}. In this work we focus on high-multiplicity of infection (MOI) single-cell CRISPR screens. We expect the ideas that we develop for this assay to apply (with some effort) to other classes of single-cell CRISPR screens as well. In this section we motivate high MOI single-cell screens, overview the experimental protocol, and present relevant analysis challenges.

The human genome consists of genes, enhancers (segments of DNA that regulate the expression of one or more genes), and other genomic elements (that are not of importance to the current work). Genome-wide association studies (GWAS) have revealed that the majority ($>90\%$) of variants associated with diseases lie outside genes and (very likely) inside enhancers \parencite{Gallagher2018}. These noncoding variants are thought to contribute to disease by modulating the expression of one or more disease-relevant genes. Scientists do not know the gene (or genes) through which most noncoding variants exert their effect, limiting the interpretability of GWAS results. A central open challenge in genetics, therefore, is to link enhancers that harbor GWAS variants to the genes that they target at genome-wide scale \parencite{Gasperini2020}.

@@

\subsection{Analysis challenges}
High MOI single-cell CRISPR screens present several statistical challenges, four of which we highlight here. Throughout, we consider a single perturbation-gene pair. First, the ``treatment'' variable -- i.e., the presence or absence of a perturbation -- cannot be directly observed. Instead, perturbed cells transcribe molecules called  \textit{guide RNAs} (or \textit{gRNAs}) that serve as indirect proxies of perturbation presence. We must leverage these gRNAs to impute (explicitly or implicitly) perturbation assignments onto the cells (Figure \ref{analysis_challenges}b). Second, ``technical factors'' -- sources of variation that are experimental rather than biological in origin -- impact the measurement of both gene and gRNA expressions and therefore act as confounders (Figure \ref{analysis_challenges}b). Third, the gene and gRNA data are sparse, discrete counts. Consequently, classical statistical approaches that assume Gaussianity or homoscedasticity are inapplicable. Finally, and most subtly, sequenced gRNAs sometimes map to cells that have not received a perturbation. This phenomenon, which we ``background contamination,'' results from errors in the sequencing and alignment processes \parencite{Replogle2020}. The marginal distribution of the gRNA counts is best conceptualized as a mixture model (Figure \ref{analysis_challenges}c; Gaussian distributions used for illustration purposes only). Unperturbed and perturbed cells both exhibit nonzero gRNA count distributions, but this distribution is shifted upward for perturbed cells. Figure \ref{analysis_challenges}d shows example data on four (of possibly tens or hundreds of thousands of) cells. The analysis objective is to leverage the gene expressions and gRNA counts to estimate the effect of the (latent) perturbation on gene expression, accounting for the technical factors.

\begin{figure}[h!]
	\centering
	\includegraphics[width=1\linewidth]{../../figures/analysis_challenges/plot.pdf}
	\caption{\textbf{Experimental design and analysis challenges}: \textbf{a,} Experimental design. For a given perturbation (e.g., the perturbation indicated in blue), we partition the cells into two groups: perturbed and unperturbed. Next, for a given gene, we conduct a differential expression analysis across the two groups, yielding an estimate of the impact of the given perturbation on the given gene. \textbf{b,} DAG representing all variables in the system. The perturbation (latent) impacts both gene expression and gRNA expression; technical factors act as confounders, also impacting gene and gRNA expression. The target of estimation is the effect of the perturbation on gene expression. \textbf{c,} Schematic illustrating the ``background read'' phenomenon. Due to errors in the sequencing and alignment processes, unperturbed cells exhibit a nonzero gRNA count distribution (bottom). The target of estimation is the change in mean gene expression in response to the perturbation (top). \textbf{d}, Example data on four cells for a given perturbation-gene pair. Note that (i) the perturbation is unobserved, and (ii) the gene and gRNA data are discrete counts.}
	\label{analysis_challenges}
\end{figure}

In this work we analyze two large-scale, high MOI, single-cell CRISPR screen datasets published by (\cite{Gasperini2019}) and (\cite{Xie2019a}). Gasperini (resp., Xie) targeted approximately 6,000 (resp., 500) candidate enhancers in a population of approximately 200,000 (resp., 100,000) cells. Gasperini additionally designed several hundred positive control, gene-targeting perturbations and 50 non-targeting, negative control perturbations to assess method sensitivity and specificity.

@@

In this work we introduced GLM-EIV (``GLM-based errors in variables''), a new method for single-cell CRISPR screen analysis. GLM-EIV extends the classical errors-in-variables model to response distributions and sources of measurement error that are exponential family-distributed and potentially impacted by the same set of confounding variables. These extensions enable GLM-EIV to resolve novel analysis challenges posed by single-cell CRISPR screens. We demonstrated through simulation studies, real data analyses, and mathematical theory that GLM-EIV outperforms thresholded regression by a considerable margin in high background contamination settings. GLM-EIV intuitively achieves this performance gain by leveraging information from \textit{both} modalities (rather than the gRNA modality alone) to assign perturbation identities to cells. On the other hand, in low background contamination settings GLM-EIV and thresholded regression work best in conjunction, with GLM-EIV providing a statistically principled and empirically effective procedure for selecting the threshold. GLM-EIV thereby neutralizes a tuning parameter that, until this point, has been selected using heuristic procedures, with little confidence that the choice is near optimal. Figure \ref{fig:dec_tree} summarizes how we anticipate GLM-EIV being used in practice.

@@

\section*{Figures}
% 1. Challenges

@@

%\clearpage
%\begin{figure}[h!]
%	\centering
%	\includegraphics[width=0.7\linewidth]{../../figures/thresholding_theoretical/plot}
%	\caption{\textbf{Thresholding method bias-variance decomposition.} Bias decreases and variance increases as the threshold tends to infinity. $\beta^g_1 = 1, \beta^m_1 = 1,$ and $\pi = 0.1$ in this plot.}
%	\label{thresholding_theoretical}
% \end{figure}

\clearpage
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.9\linewidth]{../../figures/main_text_sim/plot}
	\caption{\textbf{Simulation study}. Columns correspond to distributions (Poisson, NB with known $\theta$, NB with estimated $\theta$), and rows correspond to metrics (bias, MSE, coverage, CI width, and time). Methods are shown in different colors; GLM-EIV (red) is masked by accelerated GLM-EIV (blue) in several panels. GLM-EIV demonstrated superior statistical performance to the thresholding method on all metrics (rows 1-4). Accelerated GLM-EIV had substantially lower computational cost than ``vanilla'' GLM-EIV (bottom row) despite demonstrating identical statistical performance (rows 1-4).}
	\label{main_text_sim}
\end{figure}
